---
title: "Lab 8"
author: "Madeline Berger"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(rgdal)
library(rasterVis)
library(maps)
library(rgeos)
library(tidyverse)
library(RColorBrewer)

# view some color palettes
# display.brewer.all()
# rainbow color scheme
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) 

#setting smaller margins for plotting
par(mar=c(2,2,1,1))



# view some color palettes
# display.brewer.all()
# rainbow color scheme
cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) 

#setting smaller margins for plotting
par(mar=c(2,2,1,1))

all_threats <- raster("full_modelnv.tif")

```



```{r}

#Creating plots to visualize data

plot(all_threats,col=cols)

# add a landmap to your shapefile. the add=T argument tells R to add it to the existing plot.
# make sure you understand what the other arguments do
plot(all_threats,ext=extent(-130,-110,24,50),col=cols)
map('world',fill=T,add=T,col='gray')



plot(all_threats,col=cols,ext=extent(-121,-117,32,35),main="Cumulative Threats") # A good extent for the Santa Barbara Channel
map('world',fill=T,add=T,col='gray')

# Creating a histogram to look at value distribution

hist(all_threats,main="Cumulative Threats Frequency")

# Using cellStates to present summary statistics, in this case: mean, sd.

cellStats(all_threats,mean)
cellStats(all_threats,sd)


```

```{r}

all_spp <- raster("ca_curr_sp_rich.tif")

plot(all_spp, col=cols)



#?crop see what the crop function does

threats_crop <- crop(all_threats,all_spp) #Crop the threats layer to the same extent at species

#?resample see what the resample function does
# NOTE: the progress='text' argument is a great tool: it prints out the progress
# of a longer-running function into the console, so you can see how the operation is going

# the method='ngb' argument specifies that we want to use a nearest neighbor algorithm to resample, instead of interpolation
spp_res <- resample(all_spp,threats_crop,method='ngb',progress='text')

# We now want to "stack" the two raster files

spp_threat_stack <- stack(threats_crop,spp_res)
plot(spp_threat_stack,col=cols)




```

####Notes from class 

General 
raster comes with simple basemaps in "map" function
typing the raster name into the console will return information: where do we put it?
-dimensions tell you have any rows, columnns and cells in your raster (giant matrix)
-resolution: the resolution of each of the individual pixels
-extent: the lat long bounds 

Plotting 
Extent argument allows you to zoom into a region, the numbers after refer to lat and long (or whatever spatial reference units your data is in) 
Can change it to zoom in on certain areas (use the plot axis to estimate it)
Also can use "zoom function" to activate a locator, and clicking around will zoom to those areas 

Histogram
Can find frequency of threats using "hist", will only use a subset of the cells since the raster is so huge
Other stats using "cellstats"


Raster Calculations:
-need to overlay more data
Steps:
-import new data (Species Richness)
-resample to the same resolution
-reclassify to find the top 20% for both 
-overlay 

####Notes from class 2 - Nelson typist

Begin by reclassifying the data to only show the top 20% of cells in each layer. Then, we want to combine both layers into one to only contain "hotspots"
-Need to reclassify 0's to NA's
-Then calculate the 80th percentile and max of the data using the "quantile" and "cellStats" functions, respectively
-Use these values for both the species data and env threat layers to create layers of binary high threat/species presence



